---
title: "Example run"
output: github_document
date: "2024-10-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

title: "Example run" output: github_document date: "2024-10-08" ---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# PART A. Running DBPM for a single grid cell

This example script steps through the BIPOLE workflow but only using the inputs, obtained from Earth system model (ESM) outputs, for a single grid cell. In example 1, we show how the original code for DBPM can be used. In example 2 we show how the same workflow and a *DBPM-like* model can be implemented in mizer, which opens up the possibility to assess a wider range of marine ecosystem model structural uncertainties and to test/benchmark the performance in terms of computational time.

### Step 1 - Inputs

DBPM request inputs for a few of the ESMs to carry out climate change simulations.

```{r}

#Get ipsl monthly inputs for 3b - file below is called pp, just read in first row of historical

 load("/homevol/julia/../../rd/gem/private/fishmip_inputs/ISIMIP3b/IPSL-CM6A-LR/IPSL-CM6A-LR_historical.RData")


 inputs<- pp %>% filter(lon==15,lat==-40.5)
 
 rm(pp)
 

## OR make some up for now  based on IPSL


# Physical variables:

# tos = 13.96293 
# tob=-0.05683944
# deptho=2776.039
# 
# # BGC variables:
# # lphy = diatoms + diazotrophs (phydiat + phymisc)
# # sphy = picophytoplankton
# 
# sphy = 0.1400772
# lphy=0.05363183



```

### Step 2 - Create plankton size spectrum inputs

```{r}

# source model functions
source("/homevol/julia/R/FISHMIP/DBPM-BIOPOLE/DBPM-BIOPOLE/dbpm/dynamic_sizebased_model_functions.R")

# Get plankton size spectrum slope and intercept (already run on above, see)

slope=GetPPIntSlope(inputs$sphy,inputs$lphy,mmin=10^-14.25, mmid=10^-10.184,mmax=10^-5.25,depth=2776.039,output="slope")

intercept=GetPPIntSlope(inputs$sphy,inputs$lphy,mmin=10^-14.25, mmid=10^-10.184,mmax=10^-5.25,depth=2776.039,output="intercept")

er=getExportRatio(inputs$sphy,inputs$lphy,depth=2776.039,sst=inputs$sst)

inputs$depth<-2776.039

```

### Step 3 - Set up and run model

### 

```{r}

# if below does not run, work with pre-processed weekly inputs for a single grid  cell

inputs <- readRDS("/homevol/julia/../../rd/gem/private/fishmip_inputs/ISIMIP3b/IPSL-CM6A-LR/historical/grid_10895_IPSL-CM6A-LR_historical.rds")

## Set up params 
    params <- sizeparam(equilibrium = TRUE
                        ,dx = 0.1
                        ,xmin.consumer.u = -3
                        ,xmin.consumer.v = -3
                        ,tmax = dim(inputs$ts)[1]/48
                        ,tstepspryr  =  48
                        ,fmort.u = 0.0
                        ,fminx.u = 0
                        ,fmort.v = 0.0
                        ,fminx.v = -1
                        ,depth = inputs$depth$depth
                        ,er = inputs$ts$er
                        ,pp = inputs$ts$intercept
                        ,slope = inputs$ts$slope
                        ,sst = inputs$ts$sst
                        ,sft = inputs$ts$sbt
                        ,lat = inputs$depth$lat
                        ,lon = inputs$depth$lon
                        ,use.init = FALSE
                        ,U.initial = NA 
                        ,V.initial = NA  
                        ,W.initial = NA) 
    
    
result_set <- sizemodel(params  =params
                            ,ERSEM.det.input = FALSE
                            ,U_mat = NULL
                            ,V_mat = NULL
                            ,W_mat = NULL
                            ,temp.effect = TRUE
                            ,use.init = FALSE
                            ,burnin.len = 300*48, aggregated=T) 
  
result_set$notrun <- ifelse(any(result_set$U[150,]=="NaN")==FALSE, FALSE, TRUE)


```

### 

### Step 4 - Save required outputs for plotting and analysis

### 

```{r}

# PLOT SIZE SPECTRUM AT FINAL TIME STEP
plotsizespectrum(result_set,params)


# SUMMARISE "aggregated" output variables for FishMIP - should be wrapped into a function

# sum biomass (need to convert to g ww per m^3, across size classes) 
      # then need to convert to g C per m^2 (multiply by depth at end)
      # conversion used is: 0.0352 g C = 1 g wet weight
      # total biomass in functional groups

# save monthly
isave <- seq(from=2, to=((dim(inputs$ts)[1])+1), by = 40)
      
      TotalUbiomass <- 0.0352*apply(result_set$U[params$ref:params$Nx,isave]*params$dx*10^params$x[params$ref:params$Nx],2,sum) 
      TotalVbiomass <- 0.0352*apply(result_set$V[params$ref.det:params$Nx,isave]*params$dx*10^params$x[params$ref.det:params$Nx],2,sum) 
      TotalW <- 0.0352*result_set$W[isave]
      
      # Biomass in functional groups and size classes - these weight classes correspond to 10, 30 , 46 and 100 cm thresholds (e.g. biomass in these sizes and up)
      wcut <- c(10, 270, 1000, 10000)
      xcutref <- wcut
      
      for (i in 1:length(wcut)) xcutref[i] = (min(which(params$x >=log10(wcut[i]))))
      
      Ubiomass10plus <- 0.0352*apply(result_set$U[xcutref[1]:params$Nx,isave]*params$dx*10^params$x[xcutref[1]:params$Nx],2,sum) 
      Ubiomass270plus <- 0.0352*apply(result_set$U[xcutref[2]:params$Nx,isave]*params$dx*10^params$x[xcutref[2]:params$Nx],2,sum) 
      
      Vbiomass10plus <- 0.0352*apply(result_set$V[xcutref[1]:params$Nx,isave]*params$dx*10^params$x[xcutref[1]:params$Nx],2,sum) 
      Vbiomass270plus <- 0.0352*apply(result_set$V[xcutref[2]:params$Nx,isave]*params$dx*10^params$x[xcutref[2]:params$Nx],2,sum) 
      
      # total catches in functional groups
      
      # sum catches (already in grams per yr, across size classes) 
      # and then they need to be converted to g ww per m^2 
      
      TotalUcatch <- apply(result_set$Y.u[,isave]*params$dx,2,sum)
      TotalVcatch <- apply(result_set$Y.v[,isave]*params$dx,2,sum) 
      
      # catches in functional groups and size classes - these weight classes correspond to 10, 30 , 46 and 100 cm thresholds (e.g. biomass in these sizes and up)
      
      Ucatch10plus <- apply(result_set$Y.u[xcutref[1]:params$Nx,isave]*params$dx,2,sum) 
      Ucatch270plus <- apply(result_set$Y.u[xcutref[2]:params$Nx,isave]*params$dx,2,sum) 
      Vcatch10plus <- apply(result_set$Y.v[xcutref[1]:params$Nx,isave]*params$dx,2,sum) 
      Vcatch270plus <- apply(result_set$Y.v[xcutref[2]:params$Nx,isave]*params$dx,2,sum) 
      
      # Size spectrum slopes 
      # log 10 abundance vs. log 10 body mass, across size range beginnig of consumer spectra up to 10^4 g (where senescence kicks in)
      # could do: normalised  - plot of log (biomass density/arithmetic width of size bin) vs log size class
      # pareto - use method as in Rogers et al 2014.
      # Uslope <- Vslope <- rep(0, length(isave))
      # for (i in 1:length(isave)){
      #  Uslope[i]<-lm(log10(result_set$U[which(params$x==-3):which(params$x==3) ,isave[i]]) ~params$x[which(params$x==-3):which(params$x==3)])$coef[2] 
      #  Vslope[i]<-lm(log10(result_set$V[which(params$x==-3):which(params$x==3),isave[i]]) ~params$x[which(params$x==-3):which(params$x==3)])$coef[2] 
      # }
      
      agg <- data.frame(TotalUbiomass,Ubiomass10plus,Ubiomass270plus,TotalUcatch,Ucatch10plus,Ucatch270plus,TotalVbiomass,Vbiomass10plus,Vbiomass270plus,TotalVcatch,Vcatch10plus,Vcatch270plus,TotalW) #,Uslope,Vslope)
      rm(TotalUbiomass,Ubiomass10plus,Ubiomass270plus,TotalUcatch,Ucatch10plus,Ucatch270plus,TotalVbiomass,Vbiomass10plus,Vbiomass270plus,TotalVcatch,Vcatch10plus,Vcatch270plus, TotalW) #, Uslope,Vslope)
      
      agg$lat <- rep(params$lat,each=length(agg[,1]))
      agg$lon <- rep(params$lon,each=length(agg[,1]))
      agg$depth <- rep(params$depth,each=length(agg[,1]))
      
      # convert all biomasses and catches from g C or g WW per m^3 to per m^2   
      agg[,1:13] <- agg[,1:13] * min(agg$depth,100)
      
      output_filename <- "dbpm_output_all.rds" 
      saveRDS(agg, file=output_filename, compress=FALSE)

## Explore some plots for this historical period
      
plot(agg$TotalUbiomass,typ="l")
plot(agg$TotalUbiomass,typ="l")
plot(agg$TotalW,typ="l")

```

# PART B. Try this with mizerReef (Beese et al. 2024)

```{r}
remotes::install_github("cmbeese/mizerReef")
install.packages("remotes")
```

Set up simple DPBM model

```{r}
# Load MizerReef package
library(mizerReef)
library(mizer)
library(mizerExperimental)

# Store MizerParams object from package in local environment
group_params<-data.frame(species=c("pel_predators","ben_detritivores"),l_max=c(1000,500),a=c(0.025,0.025),b=c(3,3),k_vb=c(0.4,0.6),biomass_observed=c(100,50),biomass_cutoff=c(3,3),refuge_user=c(F,F),bad_pred=c(F,F),interaction_resource=c(1.0,0.1),interaction_algae=c(0.1,0.1),interaction_detritus=c(0.1,1),satiation=c(F,F))
inter<-rbind(c(0.5,0.5),c(0,0))
refuge<- data.frame(L_refuge = 10, prop_protect = 0.95)

## Set model ----------------------------------------
  
params<-newReefParams(group_params,interaction=inter,method="sigmoidal",method_params =refuge,degrade=FALSE)

#check interactions
params@interaction


# getting an error about degrade so overwriting
params@other_params$degrade <-FALSE
getReproductionLevel(params)

## set density dependent of reproduction ----------------
rdi <- rep(0.9, dim(group_params)[1])
params <- setBevertonHolt(params, reproduction_level = rdi)

## Project to first steady state -------------------------------
params <- params |>
      reefSteady() |> reefSteady() |> reefSteady() |> reefSteady() |>
      reefSteady() |> reefSteady()

## Calibrate biomasses and growth ---------------------------------
  # Match observed species group biomasses
  params <- calibrateReefBiomass(params)
 # params <- matchBiomasses(params)
  params <- reefSteady(params)
  

## Summary plots
   dbpm <- reefSteady(params,return_sim = T)
plotBiomass(dbpm)
plotProductivity(params)
plotRelativeContribution(params)
plotDetritusProduction(params)


## Fish the model
dbpm  <- project(params,effort=0,t_max=100)
dbpm_fishing  <- project(params,effort=0.8,t_max=100)
plotSpectra2(dbpm,dbpm_fishing)


```
