---
title: "Example run"
output: github_document
date: "2024-10-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Running DBPM for a single grid cell

This example script steps through the BIPOLE workflow but only using the inputs, obtained from Earth system model (ESM) outputs, for a single grid cell. In example 1, we show how the original code for DBPM can be used. In example 2 we show how the same workflow and a *DBPM-like* model can be implemented in mizer, which opens up the possibility to assess a wider range of marine ecosystem model structural uncertainties and to test/benchmark the performance in terms of computational time.

### Step 1 - Inputs

DBPM request inputs for a few of the ESMs to carry out climate change simulations.

```{r}

## Let's make some up for now 

# Physical variables:
tos = 20
tob=10
deptho=100

# BGC variables:
# lphy = diatoms + diazotrophs (phydiat + phymisc)
# sphy = picophytoplankton

question:
sphy = phyc - (phydiat + phymisc)

# Get plankton size spectrum slope and intercept

slope=GetPPIntSlope(sphy,lphy,mmin=10^-14.25, mmid=10^-10.184,mmax=10^-5.25,depth,output="slope")

intercept=GetPPIntSlope(sphy,lphy,mmin=10^-14.25, mmid=10^-10.184,mmax=10^-5.25,depth,output="intercept")

```

### Step 2 - Create plankton size spectrum inputs

```{r}
## Read in grid of historical (time-sliced) outputs, UK-ESM. 






```

### Step 3 - Set up and run model

### 

```{r}
## Read in grid of historical (time-sliced) outputs, UK-ESM. 






```

### 

### Step 4 - Save required outputs for plotting and analysis

### 

```{r}
## Read in grid of historical (time-sliced) outputs, UK-ESM. 






```

### Try this with mizerReef (Beese et al. 2024)

```{r}
remotes::install_github("cmbeese/mizerReef")
install.packages("remotes")
```

Set up simple DPBM model

```{r}
# Load MizerReef package
library(mizerReef)
library(mizer)
library(mizerExperimental)

# Store MizerParams object from package in local environment
group_params<-data.frame(species=c("pel_predators","ben_detritivores"),l_max=c(1000,500),a=c(0.025,0.025),b=c(3,3),k_vb=c(0.4,0.6),biomass_observed=c(100,50),biomass_cutoff=c(3,3),refuge_user=c(F,F),bad_pred=c(F,F),interaction_resource=c(1.0,0.1),interaction_algae=c(0.1,0.1),interaction_detritus=c(0.1,1),satiation=c(F,F))
inter<-rbind(c(0.5,0.5),c(0,0))
refuge<- data.frame(L_refuge = 10, prop_protect = 0.95)

## Set model ----------------------------------------
  
params<-newReefParams(group_params,interaction=inter,method="sigmoidal",method_params =refuge,degrade=FALSE)

#check interactions
params@interaction


# getting an error about degrade so overwriting
params@other_params$degrade <-FALSE
getReproductionLevel(params)

## set density dependent of reproduction ----------------
rdi <- rep(0.9, dim(group_params)[1])
params <- setBevertonHolt(params, reproduction_level = rdi)

## Project to first steady state -------------------------------
params <- params |>
      reefSteady() |> reefSteady() |> reefSteady() |> reefSteady() |>
      reefSteady() |> reefSteady()

## Calibrate biomasses and growth ---------------------------------
  # Match observed species group biomasses
  params <- calibrateReefBiomass(params)
 # params <- matchBiomasses(params)
  params <- reefSteady(params)
  

## Summary plots
   dbpm <- reefSteady(params,return_sim = T)
plotBiomass(dbpm)
plotProductivity(params)
plotRelativeContribution(params)
plotDetritusProduction(params)


## Fish the model
dbpm  <- project(params,effort=0,t_max=100)
dbpm_fishing  <- project(params,effort=0.8,t_max=100)
plotSpectra2(dbpm,dbpm_fishing)


```

