---
title: "Example run"
output: github_document
date: "2024-10-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Running DBPM for a single grid cell

This example script steps through the BIPOLE workflow but only using the inputs, obtained from Earth system model (ESM) outputs, for a single grid cell. In example 1, we show how the original code for DBPM can be used. In example 2 we show how the same workflow and a *DBPM-like* model can be implemented in mizer, which opens up the possibility to assess a wider range of marine ecosystem model structural uncertainties and to test/benchmark the performance in terms of computational time.

### Step 1 - Inputs

DBPM request inputs for a few of the ESMs to carry out climate change simulations.

```{r}

## Let's make some up for now 

# Physical variables:
tos = 20
tob=10
deptho=100

# BGC variables:
# lphy = diatoms + diazotrophs (phydiat + phymisc)
# sphy = picophytoplankton

question:
sphy = phyc - (phydiat + phymisc)

# Get plankton size spectrum slope and intercept

slope=GetPPIntSlope(sphy,lphy,mmin=10^-14.25, mmid=10^-10.184,mmax=10^-5.25,depth,output="slope")

intercept=GetPPIntSlope(sphy,lphy,mmin=10^-14.25, mmid=10^-10.184,mmax=10^-5.25,depth,output="intercept")

```

### Step 2 - Create plankton size spectrum inputs

```{r}
## Read in grid of historical (time-sliced) outputs, UK-ESM. 






```

### Step 3 - Set up and run model

### 

```{r}
## Read in grid of historical (time-sliced) outputs, UK-ESM. 






```

### 

### Step 4 - Save required outputs for plotting and analysis

### 

```{r}
## Read in grid of historical (time-sliced) outputs, UK-ESM. 






```

### 

### 
